### 2020-09-05 manuscript/ERL_global_C_review.Rmd
manuscript updates

Norbert Kunert

@@

### Summary
*(The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion. Up to 7 Keywords should also be provided. The total length of the article is flexible.)*

@@

### Background
*(The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.)*

Carbon (C) cycling in forests is central to sustaining life on Earth. Forest ecosystems globally influence climate through their vital role in regulating atmospheric carbon dioxide (CO~2~). Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ (Bagdley et al. 2019), or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. The resulting in C sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests (Le Quéré et al 2017, Schimel et al 2015) (**update, give range**). The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; Grassi et al 2017; IPCC1.5).

@@

Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age. While remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al. 2019, [Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) ([Clark et al. 2004](https://www.onlinelibrary.wiley.com/doi/abs/10.1046/j.1461-0248.2003.00547.x); [Leithold et al 2018](https://nph-onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/10.1111/nph.15110)), and to some extent net ecosystem exchange ($NEE$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements. Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements]. Thus, ground-based measurements are--and will continue to be--central to characterizing forest C cycling. Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012). Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al. 2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables (but see Anderson-Teixeira et al 2016 for the tropics) and do not consider age-by-biome interactions. To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate, boreal)?

@@

While components of these questions have been previously addressed (*e.g.*, REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.

### Methods/ Design
*(The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review. While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).)*

This review synthesizes data from the ForC database (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018). Particular attention was given to developing the database for tropical forests (Anderson-Teixeira et al 2016), yet these represented only approximately one-third of records in ForC v.2.0 (Anderson-Teixeira et al 2018). Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database (Bond-Lamberty and Thomson 2010; *hopefully Cook-Patton*; see [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)) …. We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010) and are recorded in the Koeppen field of the ForC SITES table. Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations (SYNMAP; Jung et al 2006) and recorded in the  dominant.veg field of the ForC MEAUREMENTS table. Stand age was obtained from the stand.age field of MEASUREMENTS table, and forests classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). Records for which stand age was unknown were excluded from the analysis. These groupings were defined in March 2018, at which point they covered 91% of the primary variable records for forests of known age, or 86% of total records, in ForC (v.2.0; Anderson-Teixeira et al 2018). The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). Duplicate or otherwise conflicting records (indicated in conflicts field of MEASUREMENTS table) were reconciled as follows. Replicate measurements (*i.e.*, replicates from within a single study) were averaged. Records that subsumed others— **i.e.**, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed. For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field). When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values. We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.

Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.  Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”). We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. We also removed all plots for which no stand history information had been retrieved.

@@

### Review Results/ Synthesis
*(The Review Results/Synthesis section presents the findings of the analyses.)*

@@

### Discussion 
(*The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.*)

@@

### Acknowledgements
All researchers whose data is included in ForC and this analysis.
A Smithsonian Scholarly Studies grant to KAT and HML.

### References
(*All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review. Such a matrix will be included 
as supplementary material. ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.*)

>>>>

@@

### Summary
```{r eval=FALSE, echo=FALSE}
# The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion. Up to 7 Keywords should also be provided. The total length of the article is flexible.
```

@@

### Background
```{r eval=FALSE, echo=FALSE}
# The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.
```

Carbon (C) cycling in forests is central to sustaining life on Earth. Forest ecosystems globally influence climate through their vital role in regulating atmospheric carbon dioxide (CO~2~). Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ (Bagdley et al. 2019), or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. The resulting in C sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass  (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**). The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017] (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; IPCC1.5).

@@

Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age. While remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al. 2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEE$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements. Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements]. Thus, ground-based measurements are, and will continue to be, central to characterizing forest C cycling.

Tens of thousands of ground-based forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccessible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012). Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al. 2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables [but see @anderson-teixeira_carbon_2016 for the tropics] and do not consider age-by-biome interactions. To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal)?

@@

While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018, @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.

### Methods/ Design
```{r eval=FALSE, echo=FALSE}
# The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review. While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).
```

This review synthesizes data from the ForC database [https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018]. ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. 
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018). 
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc:_2018]. Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment (GROA) database (Cook-Patton et al. in press), and [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv). 
We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations (SYNMAP; Jung et al 2006) and recorded in the  dominant.veg field of the ForC MEAUREMENTS table. 
Stand age was obtained from the stand.age field of MEASUREMENTS table, and forests classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Records for which stand age was unknown were excluded from the analysis. 
These groupings were defined based on ForC v.2.0[@anderson-teixeira_forc:_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. 
The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. 
In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX A*. 
Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.  Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”). We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. We also removed all plots for which no stand history information had been retrieved.

@@

### Review Results/ Synthesis
```{r eval=FALSE, echo=FALSE}
# The Review Results/Synthesis section presents the findings of the analyses.
```

@@

### Discussion 
```{r eval=FALSE, echo=FALSE}
# The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.
```

@@

### Acknowledgements
All researchers whose data is included in ForC and this analysis.
Thanks to Norbert Kunert for helpful discussion.
A Smithsonian Scholarly Studies grant to KAT and HML.

```{r eval=FALSE, echo=FALSE}
# All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review. Such a matrix will be included as supplementary material. ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.
```
